Introducing ICBe: Very High Recall and Precision Event Extraction from Narratives about International Crises

### 2022-07-25 replication_paper/pnas_draft/paper.Rmd

#https://bookdown.org/yihui/bookdown/a-single-document.html
#https://stackoverflow.com/questions/52531637/knitr-rmarkdown-latex-how-to-cross-reference-figures-and-tables-in-2-different/52532269#52532269
#https://stackoverflow.com/questions/51595939/bookdown-cross-reference-figure-in-another-file
#https://stackoverflow.com/questions/25824795/how-to-combine-two-rmarkdown-rmd-files-into-a-single-output/51521542#51521542
output: 
  bookdown::pdf_book:
    toc: no
    base_format: rticles::pnas_article
    latex_engine: xelatex

#Disabled several the caused errors
header-includes:
  - \usepackage{xr} \externaldocument{appendix} #External references from the appendix
#  - \usepackage{tinytex}
  - \usepackage[utf8]{inputenc}
  - \usepackage{pifont}
  - \usepackage{newunicodechar}
  - \newunicodechar{✓}{\ding{51}}
  - \newunicodechar{✗}{\ding{55}}
  - \usepackage{array}
  - \usepackage{ctable} # added for demo
  - \usepackage{natbib} #added for latex citation within huxtable
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{titlesec}
  - \usepackage[parfill]{parskip}
  - \usepackage{makecell}
  - \usepackage{graphicx}
#  - \usepackage{caption}
#  - \usepackage[capposition=top]{floatrow}
#  - \titleformat{\subsubsection}{\normalfont\normalsize\itshape}{\thesubsubsection}{1em}{}
#  - \titlespacing*{\subsubsection}{0pt}{3.25ex plus 1ex minus .2ex}{0ex plus .2ex}
#  - \DeclareUnicodeCharacter{00A0}{ }
  - \usepackage{setspace}
  - \usepackage{cellspace}
  - \setlength\cellspacetoplimit{0.8ex}
  - \renewcommand{\arraystretch}{0.8}
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
#  - \usepackage[ruled,vlined,linesnumbered]{algorithm2e}
#  - \usepackage{subfigure}
  - \usepackage{subcaption}
  - \usepackage{atbegshi}% http://ctan.org/pkg/atbegshi
  - \usepackage{float}
#  - \usepackage{eso-pic,graphicx,transparent}
#  - \usepackage[perpage]{footmisc}
  - \usepackage[algo2e]{algorithm2e}
#https://tex.stackexchange.com/questions/131646/algorithm2e-command-algorithm-already-defined
---

```{r, echo=F, cache=F, messages=F, results='hide', warnings=F, include=FALSE}
#International Crisis Behavior Events (ICBe) Dataset
knitr::opts_chunk$set(echo=F, cache=TRUE, messages=F, results='hide', warnings=F, include=FALSE) #fig.width = 8, fig.height = 8, message=F, warning=F, , results=F

@@

ggsave(file=paste0(here::here(), '/replication_paper/figures/p_recall.png'),
       plot = p_recall ,
       width=5, height=3.0)

@@

tiger <- image_read(here::here("replication_paper", "pnas_draft", "recall_cuban_and_crimea.png"))
tiger2 <- image_read(paste0(here::here(), '/replication_paper/figures/p_recall.png'))

@@

p_196_icbe <- icbe_crisismaps(thiscrisis=196) #readRDS(file=paste0(here::here(), '/replication_paper/figures/crisismaps/p_metro_plot_471.Rds')) +
    labs(
      #title = plot_title ,
      #subtitle = subtitle ,
      caption = NULL
    ) +
    theme(plot.margin=unit(c(0,0,0,0),"in")) +
   scale_x_continuous(expand = expansion(mult = c(0.025, 0.025))) +
   scale_y_continuous(expand = expansion(mult = c(0.01, 0.01)))

p_471_icbe <- icbe_crisismaps(thiscrisis=471) #readRDS(file=paste0(here::here(), '/replication_paper/figures/crisismaps/p_metro_plot_471.Rds')) +
    labs(
      #title = plot_title ,
      #subtitle = subtitle ,
      caption = NULL
    ) +
    theme(plot.margin=unit(c(0,0,0,0),"in")) +
   scale_x_continuous(expand = expansion(mult = c(0.025, 0.025))) +
   scale_y_continuous(expand = expansion(mult = c(0.01, 0.01)))

@@

save_plot(file=paste0(here::here(), '/replication_paper/figures/p_precision_combined.png'),
          p_precision_combined, base_height = 18, base_width = 24)

@@

\clearpage
\onecolumn
\begin{sidewaysfigure}[ht]
\caption{Crisis Maps  \label{fig:p_precision_combined}}
\centering{\includegraphics[width=24cm]{../figures/p_precision_combined.png}}
\end{sidewaysfigure}
\clearpage
\twocolumn

@@

ggsave(file=paste0(here::here(), '/replication_paper/figures/p_semantic_embeddings.png'), plot = p_semantic_embeddings, width=12, height=8)

@@

ggsave(file=paste0(here::here(), '/replication_paper/figures/p_semantic_embeddings_dendro.png'), plot = p_combined, width=12, height=10)

@@

\clearpage
\onecolumn
\begin{figure}[H]
\caption{ICBe event codings in comparison to Semantic Embeddings from source sentences\label{fig:semantic_embeddings}}
\centering{\includegraphics[width=17cm]{../figures/p_semantic_embeddings_dendro.png}}
\textit{Notes: Dots represent individual ICB narrative sentences, as embeded by the Paraphrase-MPNET-base-v2 large language model and flattened into two dimensions with UMAP. Text labels reflect individual leaves of the ICBe ontology, and colors represent intermediate branches of the ontology. Label placement is the median of all of the sentences that tag was applied to by the coders. The dendrogram shows hiearchical clustering of the tags. If ICBe precision is high, the sentences tags were applied to ought to say similar things, and the intended shape of the ontology ought to visually recognizable. }
\end{figure}
\clearpage
\twocolumn

Finally, how does ICBe's precision compare to the existing state of the art? The crisis-maps reveal the episode level datasets like MIDs or the original ICB are too sparse and vague to reconstruct the structure of the crisis. On the other end of the spectrum, the high recall dictionary based event datasets like Terrier and ICEWs produce so many noisy events (several hundreds thousands) that even with heavy filtering their crisis maps are completely unintelligible. Further, because of copyright issues, none of these datasets directly provide the original text spans making event level precision difficult to verify.

@@

save_plot(file=paste0(here::here(),
                      '/replication_paper/figures/p_precision_icews.png'),
          p_ICEWS, base_height = 10, base_width = 14)

@@

We investigated event extraction from narratives describing key historical episodes in international relations. We synthesized a prior belief about the latent unobserved phenomena that drive these events in international relations and proposed a mapping to observable concepts that enter into the observed historical record. We designed an ontology with high coverage over those concepts and developed a training procedure and technical stack for human coding of historical texts. Multiple validity checks find the resulting codings have high internal validity (e.g. intercoder agreement) and external validity (i.e. matching source material in both micro-details at the sentence level and macro-details spanning full historical episodes). Further, these codings perform much better in terms of recall, precision, coverage, and overall coherence in capturing these historical episodes than existing event systems used in international relations.

>>>>

#https://bookdown.org/yihui/bookdown/a-single-document.html
#https://stackoverflow.com/questions/52531637/knitr-rmarkdown-latex-how-to-cross-reference-figures-and-tables-in-2-different/52532269#52532269
#https://stackoverflow.com/questions/51595939/bookdown-cross-reference-figure-in-another-file
#https://stackoverflow.com/questions/25824795/how-to-combine-two-rmarkdown-rmd-files-into-a-single-output/51521542#51521542
#https://stackoverflow.com/questions/64700865/using-r-rstudio-knit-to-pdf-how-to-pass-custom-flags-to-the-pdflatex-command

bibliography: arxiv.bib
biblio-style: unsrt
output: rticles::arxiv_article

#Disabled several the caused errors
header-includes:
#  - \usepackage{xr} \externaldocument{appendix} #External references from the appendix
#  - \usepackage{tinytex}
  - \usepackage[utf8]{inputenc}
  - \usepackage{pifont}
  - \usepackage{newunicodechar}
  - \newunicodechar{✓}{\ding{51}}
  - \newunicodechar{✗}{\ding{55}}
  - \usepackage{array}
  - \usepackage{ctable} # added for demo
#  - \usepackage{natbib} #added for latex citation within huxtable
#  - \usepackage{biblatex} #added for latex citation within huxtable
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{titlesec}
  - \usepackage[parfill]{parskip}
  - \usepackage{makecell}
  - \usepackage{graphicx}
#  - \usepackage{caption}
#  - \usepackage[capposition=top]{floatrow}
#  - \titleformat{\subsubsection}{\normalfont\normalsize\itshape}{\thesubsubsection}{1em}{}
#  - \titlespacing*{\subsubsection}{0pt}{3.25ex plus 1ex minus .2ex}{0ex plus .2ex}
#  - \DeclareUnicodeCharacter{00A0}{ }
  - \usepackage{setspace}
  - \usepackage{cellspace}
  - \setlength\cellspacetoplimit{0.8ex}
  - \renewcommand{\arraystretch}{0.8}
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
#  - \usepackage[ruled,vlined,linesnumbered]{algorithm2e}
#  - \usepackage{subfigure}
  - \usepackage{subcaption}
  - \usepackage{atbegshi}% http://ctan.org/pkg/atbegshi
  - \usepackage{float}
#  - \usepackage{eso-pic,graphicx,transparent}
#  - \usepackage[perpage]{footmisc}
  - \usepackage[algo2e]{algorithm2e}
#  - \usepackage[backend = biber]{biblatex} #added in hopes of getting a bbl
#https://tex.stackexchange.com/questions/131646/algorithm2e-command-algorithm-already-defined
---

```{r, echo=F, cache=F, messages=F, results='hide', warnings=F, include=FALSE}
print(wd())
#International Crisis Behavior Events (ICBe) Dataset
knitr::opts_chunk$set(echo=F, cache=TRUE, messages=F, results='hide', warnings=F, include=FALSE) #fig.width = 8, fig.height = 8, message=F, warning=F, , results=F

#Execute these lines to get bbl files for the arxiv submission
options(tinytex.clean=F)

@@

ggsave(file=paste0(here::here(), '/replication_paper/pnas_draft/p_recall.png'),
       plot = p_recall ,
       width=5, height=3.0)

@@

tiger <- image_read(here::here("replication_paper", "pnas_draft", "recall_cuban_and_crimea.png"))
tiger2 <- image_read(paste0(here::here(), '/replication_paper/pnas_draft/p_recall.png'))

@@

p_196_icbe <- icbe_crisismaps(thiscrisis=196) #readRDS(file=paste0(here::here(), '/replication_paper/pnas_draft/crisismaps/p_metro_plot_471.Rds')) +
    labs(
      #title = plot_title ,
      #subtitle = subtitle ,
      caption = NULL
    ) +
    theme(plot.margin=unit(c(0,0,0,0),"in")) +
   scale_x_continuous(expand = expansion(mult = c(0.025, 0.025))) +
   scale_y_continuous(expand = expansion(mult = c(0.01, 0.01)))

p_471_icbe <- icbe_crisismaps(thiscrisis=471) #readRDS(file=paste0(here::here(), '/replication_paper/pnas_draft/crisismaps/p_metro_plot_471.Rds')) +
    labs(
      #title = plot_title ,
      #subtitle = subtitle ,
      caption = NULL
    ) +
    theme(plot.margin=unit(c(0,0,0,0),"in")) +
   scale_x_continuous(expand = expansion(mult = c(0.025, 0.025))) +
   scale_y_continuous(expand = expansion(mult = c(0.01, 0.01)))

@@

ratio=1.5
height=18
save_plot(file=paste0(here::here(), '/replication_paper/pnas_draft/p_precision_combined.png'),
          p_precision_combined, base_height = 18, base_width = height*ratio)

@@

\clearpage
\onecolumn
\begin{sidewaysfigure}[ht]
\caption{Crisis Maps  \label{fig:p_precision_combined}}
\centering{\includegraphics[width=24cm]{./p_precision_combined.png}}
\end{sidewaysfigure}
\clearpage
\twocolumn

@@

ggsave(file=paste0(here::here(), '/replication_paper/pnas_draft/p_semantic_embeddings.png'), plot = p_semantic_embeddings, width=12, height=8)

@@

ggsave(file=paste0(here::here(), '/replication_paper/pnas_draft/p_semantic_embeddings_dendro.png'), plot = p_combined, width=12, height=10)

@@

\clearpage
\onecolumn
\begin{figure}[H]
\caption{ICBe event codings in comparison to Semantic Embeddings from source sentences\label{fig:semantic_embeddings}}
\centering{\includegraphics[width=17cm]{./p_semantic_embeddings_dendro.png}}
\textit{Notes: Dots represent individual ICB narrative sentences, as embeded by the Paraphrase-MPNET-base-v2 large language model and flattened into two dimensions with UMAP. Text labels reflect individual leaves of the ICBe ontology, and colors represent intermediate branches of the ontology. Label placement is the median of all of the sentences that tag was applied to by the coders. The dendrogram shows hiearchical clustering of the tags. If ICBe precision is high, the sentences tags were applied to ought to say similar things, and the intended shape of the ontology ought to visually recognizable. }
\end{figure}
\clearpage
\twocolumn

Finally, how does ICBe's precision compare to the existing state of the art? The crisis-maps reveal the episode level datasets like MIDs or the original ICB are too sparse and vague to reconstruct the structure of the crisis (SI Appendix Figure D3 and D4). On the other end of the spectrum, the high recall dictionary based event datasets like Terrier and ICEWs produce so many noisy events (several hundreds thousands) that even with heavy filtering their crisis maps are completely unintelligible. Further, because of copyright issues, none of these datasets directly provide the original text spans making event level precision difficult to verify.

@@

save_plot(file=paste0(here::here(),
                      '/replication_paper/pnas_draft/p_precision_icews.png'),
          p_ICEWS, base_height = 10, base_width = 14)

@@

We investigated event abstraction from narratives describing key historical episodes in international relations. We synthesized a prior belief about the latent unobserved phenomena that drive these events in international relations and proposed a mapping to observable concepts that enter into the observed historical record. We designed an ontology with high coverage over those concepts and developed a training procedure and technical stack for human coding of historical texts. Multiple validity checks find the resulting codings have high internal validity (e.g. intercoder agreement) and external validity (i.e. matching source material in both micro-details at the sentence level and macro-details spanning full historical episodes). Further, these codings perform much better in terms of recall, precision, coverage, and overall coherence in capturing these historical episodes than existing event systems used in international relations.
